Estimating autocorrelation from multiple Markov chains

### 2020-06-12 Paper/main.tex
Further organizing the appendix

\begin{align*}
    \tilde{\Sigma}_s &= \sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\tilde{\Gamma}_s(k)\\
    \tilde{\Gamma}_s(k) &= \dfrac{1}{n}\sum_{t=1}^{n-|k|}(X_{st}-\mu)(X_{s(t+k)}-\mu)^T
\end{align*}

The average pseudo spectral variance estimator (APSVE) is given by

\[
\tilde{\Sigma} = \dfrac{1}{m}\sum\limits_{s=1}^{m}\tilde{\Sigma}_s
\]

@@

\begin{figure}[h]
    \centering
    \includegraphics[width = .7\textwidth]{plots/boom-ess.pdf}
    \caption{Running plot for $\hat{ESS}/mn$ using ASV and RSV calculated for two chains (left) and five chains (right). The parameters for target distribution are $A == 1, \; B = 3, \; C = 8.$}
    \label{fig:run_ess}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width = .7\textwidth]{plots/boom-ess_1_10_7.pdf}
    \caption{Running plot for $\hat{ESS}/mn$ using ASV and RSV calculated for two chains (left) and five chains (right). The parameters of target distribution are $A = 1, \; B = 10, \; C = 7$}
    \label{fig:boom-ess_1_10_7}
\end{figure}

@@

\begin{lemma} \label{lemma:consis_1}
  Let the strong invariance principle hold. If $\dfrac{\psi(n)}{n} \to 0$ as $n \to \infty$, then $\| \bar{\bar{X}} - \mu\|_{\infty}, \|\bar{X}_s - \mu\|_{\infty} \xrightarrow[]{a.s.} 0$ as $n \to \infty$ for all $s \in \{1,..., m\}$\\
  \end{lemma}

@@

Using the results of \cite{song1995optimal} given in \eqref{eq:song} to expand $\Var(\bar{X}_1)$,
\[
\mathbb{E}\left[\hat{\Gamma}_{RAC}(k)\right] = \left(1- \dfrac{|k|}{n}\right)\left(\Gamma(k) - \dfrac{\Sigma}{mn} - \dfrac{\Phi}{mn^2}\right)  + \dfrac{2(m-1)}{m}\dfrac{\abs{k}}{n^2} \left(\Sigma - \sum_{h=0}^{n-1}\Gamma(h)\right) + o(n^{-2})\,.
\]
%
We observe that this expectation result is very similar to equation \ref{eq:priestly} from \cite{priestley1981spectral} for the asymptotically unbiased estimator of autocovariance except for an additional positive bias term. This helps in countering the inherent negative bias in $\hat{\Gamma}(k)$ and give more informative estimates of autocovariance.  Although, bias term for a fixed $k$ is $o(1)$, it increases with $k$. We have explicitly written the end effect term arising from using the estimator $\bar{\bar{X}}$ for $\mu$. These are small order terms that vanish as $n \to \infty$.
\end{proof}

@@

\begin{align*}
    \hat{\Gamma}_{RAC}(0) &= \dfrac{1}{m}\sum_{s=1}^{m}\hat{\Gamma}_{RAC,s}(0) = \dfrac{1}{mn}\sum_{s=1}^{m}\sum_{t=1}^{n}  \left(X_{st} - \bar{\bar{X}}\right)  \left(X_{st} - \bar{\bar{X}}\right)^T\\
    &= \left[\dfrac{1}{mn}\sum_{s=1}^{m}\sum_{t=1}^{n} \left(X_{st} - \bar{X}_s \right)  \left(X_{st} - \bar{X}_s \right)^T\right] + \left[\dfrac{1}{m}\sum_{s=1}^{m} \left(\bar{X}_s - \bar{\bar{X}} \right) \left(\bar{X}_s - \bar{\bar{X}} \right)^T\right]\\
    &= \dfrac{1}{m}\sum_{s=1}^{m}\hat{\Gamma}_s(k)  + \dfrac{1}{m}\sum_{s=1}^{m} \left(\bar{X}_s - \bar{\bar{X}} \right)  \left(\bar{X}_s - \bar{\bar{X}} \right)^T\,.
\end{align*}
Using \eqref{eq:priestly}, Lemma~\ref{lemma:bias2}, and \eqref{eq:song} the expectation of $\hat{\Gamma}_{RAC}(0)$ can be written as 
%
\begin{align*}
    \mathbb{E} \left[\hat{\Gamma}_{RAC}(0) \right] &= \left(\Gamma(k) - \Var(\bar{X}_1)\right) + \left(1-\dfrac{1}{m}\right)\Var(\bar{X}_1) = \Gamma(k) - \dfrac{1}{m}\Var(\bar{X}_1)\\
    &= \Gamma(0) - \dfrac{\Sigma}{mn} - \dfrac{\Phi}{mn^2} + o(n^{-2})
\end{align*}

@@

\begin{lemma} \label{lemma:rsv_breakdown}
    The RSV estimator can be written as the sum of average pseudo SVE and two terms given by:
    \[
    \hat{\Sigma}_{RSV}^{ij} = \tilde{\Sigma}^{ij} + M_1 + M_2\,,
    \]
    where 
\begin{align*}
  M_1 & = \dfrac{1}{m}\sum\limits_{s=1}^{m}\left\{\sum\limits_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\sum\limits_{t=1}^{n-|k|}\dfrac{1}{n}\left[ \left(X_{st}-\mu \right)_i   \left(\mu-\bar{\bar{X}} \right)_j +    \left(\mu-\bar{\bar{X}} \right)_i  \left(X_{s(t+k)}-\mu \right)_j \right]\right\}\,, \\ 
  %
M_2 &= \left(\mu-\bar{\bar{X}} \right)_i   \left(\mu-\bar{\bar{X}} \right)_j\sum\limits_{k=-b_n+1}^{b_n-1}\left(1-\dfrac{|k|}{n}\right)w\left(\dfrac{k}{n}\right)\,.
\end{align*}
\end{lemma}

\begin{proof}
The proof follows from standard algebraic calculations and is presented here for completeness. Consider,
\begin{align*}
&\hat{\Sigma}_{RSV}^{ij} \\
&= \dfrac{1}{m}\sum_{s=1}^{m} \sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\dfrac{1}{n}\sum_{t=1}^{n-|k|} \left(X_{st}-\bar{\bar{X}} \right)_i \left(X_{s(t+k)}-\bar{\bar{X}} \right)_j\\
&= \dfrac{1}{m}\sum_{s=1}^{m}\sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\dfrac{1}{n}\sum_{t=1}^{n-|k|} \left[  \left\{\left(X_{st}-\mu \right)_i + \left(\mu-\bar{\bar{X}} \right)_i \right \}  \left\{ \left(X_{s(t+k)}-\mu \right)_j + \left(\mu-\bar{\bar{X}} \right)_j \right \} \right]\\
&= \dfrac{1}{m}\sum_{s=1}^{m}\sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\dfrac{1}{n}\sum_{t=1}^{n-|k|}  \left[ \left(X_{st}-\mu \right)_i  \left(X_{s(t+k)}-\mu \right)_j+  \left(X_{st} - \mu \right)_i    \left(\mu - \bar{\bar{X}} \right)_j \right. \\  
& \quad + \left. \left(\mu-\bar{\bar{X}} \right)_i  \left(X_{s(t+k)}-\mu \right)_j + \left(\mu-\bar{\bar{X}} \right)_i  \left(\mu-\bar{\bar{X}}  \right)_j  \right]\\
& = \tilde{\Sigma}^{ij} + \left[(\mu-\bar{\bar{X}})_i(\mu-\bar{\bar{X}})_j\sum_{k=-b_n+1}^{b_n-1}\left(1-\dfrac{|k|}{n}\right)w\left(\dfrac{k}{n}\right)\right] \\ 
& \quad  + \dfrac{1}{m}\sum_{s=1}^{m}\sum_{k=-b_n+1}^{b_n-1}  w\left(\dfrac{k}{b_n}\right)\sum_{t=1}^{n-|k|}  \left[\dfrac{1}{n} \left(X_{st} - \mu \right)_i \left(\mu - \bar{\bar{X}} \right)_j + \dfrac{1}{n} \left(\mu-\bar{\bar{X}} \right)_i  \left(X_{s(t+k)}-\mu \right)_j \right] \\ 
& = \tilde{\Sigma}^{ij} + M_1 + M_2\,.
\end{align*}
%
% Therefore we get the following two terms:\\\\
% Term (1): $\dfrac{1}{m}\sum\limits_{s=1}^{m}\left\{\sum\limits_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\sum\limits_{t=1}^{n-|k|}\left[\dfrac{1}{n}\left((X_{st}-\mu)_i(\mu-\bar{\bar{X}})_j\right)+ \dfrac{1}{n}\left((\mu-\bar{\bar{X}})_i(X_{s(t+k)}-\mu)_j\right) \right]\right\}$\\
% Term (2): $(\mu-\bar{\bar{X}})_i(\mu-\bar{\bar{X}})_j\sum\limits_{k=-b_n+1}^{b_n-1}\left(1-\dfrac{|k|}{n}\right)w\left(\dfrac{k}{n}\right)$\\
% \\\\
\end{proof}

\begin{lemma} \label{lemma:pseudo_consis}
If the Assumption \ref{ass:sip}, \ref{ass:sve_consis} hold and $n^{-1}{b_n \log \log n}\to 0$, then $\tilde{\Sigma}^{ij} \xrightarrow{a.s.} \Sigma^{ij} \textrm{ as } n \to \infty$
\end{lemma}

\begin{proof}
We have,
\begin{align*}
    &\tilde{\Sigma}^{ij} \\
    & = \dfrac{1}{m}\sum_{s=1}^{m}\sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\dfrac{1}{n}\sum_{t=1}^{n-|k|}  \left(X_{st} - \mu \right)_i  \left(X_{s(t+|k|)} - \mu \right)_j\\
    & = \dfrac{1}{m}\sum_{s=1}^{m}\sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\dfrac{1}{n}\sum_{t=1}^{n-|k|}  \left(X_{st} \pm \bar{X}_s - \mu \right)_i  \left(X_{s(t+|k|)} \pm \bar{X}_s - \mu \right)_j\\
    & = \hat{\Sigma}_{SV}^{ij} + \dfrac{1}{m}\sum_{s=1}^{m}\sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\dfrac{1}{n}\sum_{t=1}^{n-|k|}\left[ \left(X_{st} - \bar{X}_s \right)_i   \left(\bar{X}_s - \mu \right)_j + \left(\bar{X}_s - \mu \right)_i  \left(X_{s(t+k)} - \bar{X}_s \right)_j\right]\\
    & \quad + \left[\dfrac{1}{m}  \sum_{s=1}^{m}  \left(\bar{X}_s - \mu \right)_i  \left(\bar{X}_s - \mu \right)_j \right]  \left[\sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\left(1 - \dfrac{\abs{k}}{b_n}\right) \right]\\
    & = \hat{\Sigma}_{SV}^{ij} + N_1 + N_2 \,,
\end{align*}
where
\begin{align*}
N_1 & = \dfrac{1}{m}\sum_{s=1}^{m}  \sum_{k=-b_n+1}^{b_n-1}  w\left(\dfrac{k}{b_n}\right)\dfrac{1}{n}  \sum_{t=1}^{n-|k|}  \left[ \left(X_{st} - \bar{X}_s \right)_i  \left(\bar{X}_s - \mu \right)_j + \left(\bar{X}_s - \mu \right)_i  \left(X_{s(t+k)} - \bar{X}_s \right)_j\right]\\
N_2 & = \left[\dfrac{1}{m}  \sum_{s=1}^{m}  \left(\bar{X}_s - \mu \right)_i  \left(\bar{X}_s - \mu \right)_j \right]  \left[\sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\left(1 - \dfrac{\abs{k}}{b_n}\right) \right]\,.
\end{align*}
% Let Term (1*) be $\dfrac{1}{m}\sum\limits_{s=1}^{m}\sum\limits_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\dfrac{1}{n}\sum\limits_{t=1}^{n-|k|}\left[(X_{st} - \bar{X}_s)_i(\bar{X}_s - \mu)_j + (\bar{X}_s - \mu)_i(X_{s(t+k)} - \bar{X}_s)_j\right]$
% Let Term (2*) be $\dfrac{1}{m}\sum\limits_{s=1}^{m}(\bar{X}_s - \mu)_i(\bar{X}_s - \mu)_j\sum\limits_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\left(1 - \dfrac{\abs{k}}{b_n}\right)$
% \\\\
We will prove that $N_1, N_2\xrightarrow{a.s.} 0 \textrm{ as } n \to \infty$. From \cite{vats:fleg:jon:2018}, we have that $\hat{\Sigma}_{SV} \to 0$ with probability 1 as $n \to \infty$ for multivariate case. Assuming that all the assumptions made there hold true for our case, we will prove that $\tilde{\Sigma} \to 0$ with probability 1 as $n \to \infty$.

Firstly, we prove that  $N_1 \xrightarrow{a.s.} 0$ as $n \to \infty$. We will prove this for the first part of $N_1$. It will hold true equivalently for the second part as well. The first part of the $N_1$ is,

\begin{align*}
    & \dfrac{1}{m}\sum_{s=1}^{m}\sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\dfrac{1}{n}\sum_{t=1}^{n-|k|}\left[ \left(X_{st} - \bar{X}_s \right)_i  \left(\bar{X}_s - \mu \right)_j \right]\\
    & = \dfrac{1}{m}\sum_{s=1}^{m}\sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\dfrac{  \left(\bar{X}_s - \mu \right)_j}{n} \left[\sum\limits_{t=1}^{|k|} \left(\bar{X}_s - X_{st} \right)_i\right]\\
    & = \dfrac{1}{m}\sum_{s=1}^{m}\sum_{k=-b_n+1}^{b_n-1} w\left(\dfrac{k}{b_n}\right) \dfrac{\left(\bar{X}_s - \mu \right)_j}{n} \left[\sum_{t=1}^{|k|} \left(\mu - X_{st} \right)_i + \abs{k} \left(\bar{X}_s - \mu \right)_i\right]\\
    & \leq \dfrac{1}{m}\sum_{s=1}^{m}\sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\dfrac{\|\bar{X}_s - \mu\|_{\infty}}{n} \left\|\sum_{t=1}^{|k|} \left(\mu - X_{st} \right) + \abs{k} \left(\bar{X}_s - \mu \right)\right\|_{\infty}\\
    & \leq \dfrac{1}{m}\sum_{s=1}^{m}\sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\dfrac{\|\bar{X}_s - \mu\|_{\infty}}{n} \left(\left\|\sum_{t=1}^{|k|}  \left(X_{st} - \mu \right)\right\|_{\infty} + \abs{k}\|\bar{X}_s - \mu\|_{\infty} \right)\\
    & \leq \dfrac{1}{m}\sum_{s=1}^{m}\sum_{k=-b_n+1}^{b_n-1}\dfrac{\|\bar{X}_s - \mu\|_{\infty}}{n}\left\|\sum_{t=1}^{|k|} \left(\bar{X}_s - \mu\right) \right\| + \dfrac{1}{m}\sum\limits_{s=1}^{m} \dfrac{b_n(b_n-1)}{n} \left \|\bar{X}_s - \mu \right \|_{\infty}^2\,.
\end{align*}

Using SIP on the summation of k terms,
\begin{align*}
    & \dfrac{1}{m}\sum_{s=1}^{m}\sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\dfrac{1}{n}\sum_{t=1}^{n-|k|}\left[ \left(X_{st} - \bar{X}_s \right)_i  \left(\bar{X}_s - \mu \right)_j \right]\\
   &  < \dfrac{1}{m}\sum\limits_{s=1}^{m}\|\bar{X}_s - \mu\|_{\infty}\sum\limits_{k=-b_n + 1}^{b_n-1}\left[ \dfrac{D \psi(k)}{n} + \dfrac{\|L\| p^{1/2}(1+\epsilon)\sqrt{2k \log\log k}}{n}  \right] + \dfrac{1}{m}\sum\limits_{s=1}^{m} \dfrac{b_n(b_n-1)}{n}\|\bar{X}_s - \mu\|_{\infty}^2\\
   &  < \dfrac{(2b_n-1)}{m}\sum\limits_{s=1}^{m}\|\bar{X}_s - \mu\|_{\infty}\left[ \dfrac{D \psi(n)}{n} + \dfrac{\|L\| p^{1/2}(1+\epsilon)\sqrt{2n \log\log n}}{n}  \right] + \dfrac{1}{m}\sum\limits_{s=1}^{m} \dfrac{b_n(b_n-1)}{n}\|\bar{X}_s - \mu\|_{\infty}^2
\end{align*}

>>>>

@@

\begin{figure}[htbp]
    \centering
    \includegraphics[width = .7\textwidth]{plots/boom-ess.pdf}
    \caption{Running plot for $\hat{ESS}/mn$ using ASV and RSV calculated for two chains (left) and five chains (right). The parameters for target distribution are $A == 1, \; B = 3, \; C = 8.$}
    \label{fig:run_ess}
\end{figure}

@@

\begin{figure}[htbp]
    \centering
    \includegraphics[width = .7\textwidth]{plots/boom-ess_1_10_7.pdf}
    \caption{Running plot for $\hat{ESS}/mn$ using ASV and RSV calculated for two chains (left) and five chains (right). The parameters of target distribution are $A = 1, \; B = 10, \; C = 7$}
    \label{fig:boom-ess_1_10_7}
\end{figure}

@@

\begin{lemma} \label{lemma:consis_1}
  If Assumption~\ref{ass:sip} holds, then $\| \bar{\bar{X}} - \mu\|_{\infty}, \|\bar{X}_s - \mu\|_{\infty} \xrightarrow[]{a.s.} 0$ as $n \to \infty$ for all $s \in \{1,..., m\}$\\
  \end{lemma}

@@

% Using the results of \cite{song1995optimal} given in \eqref{eq:song} to expand $\Var(\bar{X}_1)$,
% \[
% \mathbb{E}\left[\hat{\Gamma}_{RAC}(k)\right] = \left(1- \dfrac{|k|}{n}\right)\left(\Gamma(k) - \dfrac{\Sigma}{mn} - \dfrac{\Phi}{mn^2}\right)  + \dfrac{2(m-1)}{m}\dfrac{\abs{k}}{n^2} \left(\Sigma - \sum_{h=0}^{n-1}\Gamma(h)\right) + o(n^{-2})\,.
% \]
%
% We observe that this expectation result is very similar to equation \ref{eq:priestly} from \cite{priestley1981spectral} for the asymptotically unbiased estimator of autocovariance except for an additional positive bias term. This helps in countering the inherent negative bias in $\hat{\Gamma}(k)$ and give more informative estimates of autocovariance.  Although, bias term for a fixed $k$ is $o(1)$, it increases with $k$. We have explicitly written the end effect term arising from using the estimator $\bar{\bar{X}}$ for $\mu$. These are small order terms that vanish as $n \to \infty$.

For a diagonal element of $\Gamma$,
\begin{align*}
& \E \left[ \hat{\Gamma}_{RAC,s}^{ii}\right]\\
& = \mathbb{E}\left[\hat{\Gamma}_{s}^{ii}(k)\right] - {\color{red}\dfrac{2|k|}{n}\dfrac{m-1}{mn}\left(\sum_{h=0}^{n-1}\Gamma^{ii}(h) - n \Var(\bar{X}_1)^{ii}\right)} + \left(1- \dfrac{|k|}{n}\right)\left(1-\dfrac{1}{m}\right)\Var(\bar{X}_1)^{ii}\,.
\end{align*}
In the presence of positive correlation, the leftover term is positive. 
\end{proof}

@@

\begin{align*}
    \hat{\Gamma}_{RAC}(0) &= \dfrac{1}{m}\sum_{s=1}^{m}\hat{\Gamma}_{RAC,s}(0) = \dfrac{1}{mn}\sum_{s=1}^{m}\sum_{t=1}^{n}  \left(X_{st} - \bar{\bar{X}}\right)  \left(X_{st} - \bar{\bar{X}}\right)^T\\
    &= \left[\dfrac{1}{mn}\sum_{s=1}^{m}\sum_{t=1}^{n} \left(X_{st} - \bar{X}_s \right)  \left(X_{st} - \bar{X}_s \right)^T\right] + \left[\dfrac{1}{m}\sum_{s=1}^{m} \left(\bar{X}_s - \bar{\bar{X}} \right) \left(\bar{X}_s - \bar{\bar{X}} \right)^T\right]\\
    &= \dfrac{1}{m}\sum_{s=1}^{m}\hat{\Gamma}_s(0)  + \dfrac{1}{m}\sum_{s=1}^{m} \left(\bar{X}_s - \bar{\bar{X}} \right)  \left(\bar{X}_s - \bar{\bar{X}} \right)^T\,.
\end{align*}
Using \eqref{eq:priestly} and \eqref{eq:acf_4} the expectation of $\hat{\Gamma}_{RAC}(0)$ can be written as 
%
\begin{align*}
    \mathbb{E} \left[\hat{\Gamma}_{RAC}(0) \right] &= \E\left[\hat{\Gamma}_1(0) \right] + \left(1-\dfrac{1}{m}\right)\Var(\bar{X}_1) \\
    & = \Gamma(k) - \dfrac{1}{m}\Var(\bar{X}_1)\,.
    % &= \Gamma(0) - \dfrac{\Sigma}{mn} - \dfrac{\Phi}{mn^2} + o(n^{-2})
\end{align*}

@@

\begin{align*}
    \tilde{\Gamma}_s(k) &= \dfrac{1}{n}\sum_{t=1}^{n-|k|}(X_{st}-\mu)(X_{s(t+k)}-\mu)^T \\ 
    \tilde{\Sigma}_s &= \sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\tilde{\Gamma}_s(k) \,.
\end{align*}
%
The average pseudo spectral variance estimator (APSVE) is given by
\[
\tilde{\Sigma} = \dfrac{1}{m}\sum\limits_{s=1}^{m}\tilde{\Sigma}_s
\]

Further, let
\begin{align*}
  M_1 & = \dfrac{1}{m}\sum\limits_{s=1}^{m}\left\{\sum\limits_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\sum\limits_{t=1}^{n-|k|}\dfrac{1}{n}\left[ \left(X_{st}-\mu \right)_i   \left(\mu-\bar{\bar{X}} \right)_j +    \left(\mu-\bar{\bar{X}} \right)_i  \left(X_{s(t+k)}-\mu \right)_j \right]\right\}\,, \\ 
  %
M_2 &= \left(\mu-\bar{\bar{X}} \right)_i   \left(\mu-\bar{\bar{X}} \right)_j\sum\limits_{k=-b_n+1}^{b_n-1}\left(1-\dfrac{|k|}{n}\right)w\left(\dfrac{k}{b_n}\right)\,.
\end{align*}
\begin{lemma} \label{lemma:rsv_breakdown}
For the RSV estimator $\hat{\Sigma}_{RSV}^{ij} = \tilde{\Sigma}^{ij} + M_1 + M_2$ and 
\[
|M_1 + M_2| \leq D^2 g_1(n) + D g_2(n) + g_3(n)\,
\]
where
\begin{align*}
    g_1(n) &= (4+C_1)\dfrac{b_n \psi^2(n)}{n^2} - 4\dfrac{\psi^2(n)}{n^2} \to 0\\
    g_2(n) &= 2\sqrt{2}\|L\|p^{1/2}(1+\epsilon)\left[(4+C_1)\dfrac{b_n\psi(n)\sqrt{n\log \log n}}{n^2} - 4\dfrac{\psi(n)\sqrt{n\log \log n}}{n^2}\right] \to 0\\
    g_3(n) &= \|L\|^2 p (1+\epsilon)^2\left[(4+C_1)\dfrac{b_n \log\log n}{n} - 4 \dfrac{\log \log n}{n}\right] \to 0\,.
\end{align*}
\end{lemma}

\begin{proof}
The proof follows from standard algebraic calculations and is presented here for completeness. Consider,
\begin{align*}
&\hat{\Sigma}_{RSV}^{ij} \\
&= \dfrac{1}{m}\sum_{s=1}^{m} \sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\dfrac{1}{n}\sum_{t=1}^{n-|k|} \left(X_{st}-\bar{\bar{X}} \right)_i \left(X_{s(t+k)}-\bar{\bar{X}} \right)_j\\
&= \dfrac{1}{m}\sum_{s=1}^{m}\sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\dfrac{1}{n}\sum_{t=1}^{n-|k|} \left[  \left\{\left(X_{st}-\mu \right)_i + \left(\mu-\bar{\bar{X}} \right)_i \right \}  \left\{ \left(X_{s(t+k)}-\mu \right)_j + \left(\mu-\bar{\bar{X}} \right)_j \right \} \right]\\
&= \dfrac{1}{m}\sum_{s=1}^{m}\sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\dfrac{1}{n}\sum_{t=1}^{n-|k|}  \left[ \left(X_{st}-\mu \right)_i  \left(X_{s(t+k)}-\mu \right)_j+  \left(X_{st} - \mu \right)_i    \left(\mu - \bar{\bar{X}} \right)_j \right. \\  
& \quad + \left. \left(\mu-\bar{\bar{X}} \right)_i  \left(X_{s(t+k)}-\mu \right)_j + \left(\mu-\bar{\bar{X}} \right)_i  \left(\mu-\bar{\bar{X}}  \right)_j  \right]\\
& = \tilde{\Sigma}^{ij} + \left[(\mu-\bar{\bar{X}})_i(\mu-\bar{\bar{X}})_j\sum_{k=-b_n+1}^{b_n-1}\left(1-\dfrac{|k|}{n}\right)w\left(\dfrac{k}{b_n}\right)\right] \\ 
& \quad  + \dfrac{1}{m}\sum_{s=1}^{m}\sum_{k=-b_n+1}^{b_n-1}  w\left(\dfrac{k}{b_n}\right)\sum_{t=1}^{n-|k|}  \left[\dfrac{1}{n} \left(X_{st} - \mu \right)_i \left(\mu - \bar{\bar{X}} \right)_j + \dfrac{1}{n} \left(\mu-\bar{\bar{X}} \right)_i  \left(X_{s(t+k)}-\mu \right)_j \right] \\ 
& = \tilde{\Sigma}^{ij} + M_1 + M_2\,.
\end{align*}

Consequently
\[
\left|\hat{\Sigma}_{RSV}^{ij} - \tilde{\Sigma}^{ij}  \right| = |M_1 + M_2| \leq |M_1| + |M_2|\,.
\]
Consider,
\begin{align*}
& |M_1| \\ 
   & = \left|\dfrac{1}{m}\sum _{s=1}^{m}\left\{\sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\sum_{t=1}^{n-|k|} \dfrac{1}{n}\left[ \left(X_{st} - \mu \right)_i  \left(\mu-\bar{\bar{X}} \right)_j + \left(\mu-\bar{\bar{X}} \right)_i  \left(X_{s(t+k)} - \mu \right)_j\right]\right\}\right| \\
   & \leq \dfrac{1}{m}\sum_{s=1}^{m}\left\{\sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\left[ \dfrac{1}{n}{\left|\sum\limits_{t=1}^{n-|k|} \left(X_{st} - \mu \right)_i  \left(\mu - \bar{\bar{X}} \right)_j\right|} + \dfrac{1}{n}{\left|\sum_{t=1}^{n-|k|} \left(\mu-\bar{\bar{X}} \right)_i\left(X_{s(t+k)} - \mu \right)_j\right|} \right]\right\}\\
    & \leq \dfrac{1}{m}\sum_{s=1}^{m}\left\{\sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\left[ \dfrac{1}{n}\left|\sum_{t=1}^{n-|k|}(X_{st}- \mu)_i\right|\left|(\mu-\bar{\bar{X}})_j\right|+ \dfrac{1}{n}\left|(\mu-\bar{\bar{X}})_i\right|\left|\sum_{t=1}^{n-|k|}(X_{j(t+k)}-\mu)_j\right|\right]\right\}\\
    & \leq \dfrac{\|(\bar{\bar{X}} - \mu)\|_{\infty}}{m} \sum_{s=1}^{m}\sum\limits_{k=-b_n+1}^{b_n-1}\left[ \dfrac{1}{n}\left\|\sum_{t=1}^{n-|k|}(X_{st}-\mu)\right\|_{\infty} + \dfrac{1}{n}\left\|\sum_{t=1}^{n-|k|}(X_{s(t+k)}-\mu)\right\|_{\infty} \right]\\
    &\leq \dfrac{\|(\bar{\bar{X}} - \mu)\|_{\infty}}{m} \\
    & \quad \times \sum_{s=1}^{m}\sum_{k=-b_n+1}^{b_n-1}\left[ \dfrac{1}{n}\left\|\sum_{t=n-|k|+1}^{n}(X_{st} - \mu) - n(\bar{X}_s - \mu) \right\|_{\infty} + \dfrac{1}{n}\left\|\sum_{t=1}^{|k|}(X_{st} - \mu) - n(\bar{X}_s - \mu)\right\|_{\infty} \right]\\
    &\leq \dfrac{\|(\bar{\bar{X}} - \mu)\|_{\infty}}{m} \sum_{s=1}^{m}\sum\limits_{k=-b_n+1}^{b_n-1}\left[ \dfrac{1}{n}\left\|\sum_{t=n-|k|+1}^{n}(X_{st} - \mu)\right\|_{\infty} + \dfrac{1}{n}\left\|\sum_{t=1}^{|k|}(X_{st} - \mu)\right\|_{\infty} + 2\|\bar{X}_s - \mu\|_{\infty} \right]\\
    & \leq \dfrac{\|(\bar{\bar{X}} - \mu)\|_{\infty}}{m} \sum\limits_{s=1}^{m}\sum_{k=-b_n + 1}^{b_n-1}   \dfrac{1}{n}\left[\left\|\sum_{t=n-|k|+1}^{n}(X_{st} - \mu)\right\|_{\infty} + \left\|\sum_{t=1}^{|k|}(X_{st} - \mu)\right\|_{\infty} \right]\\
    & \; \;+ 2(2b_n - 1)\|\bar{\bar{X}} - \mu\|_{\infty}\|\bar{X}_1 - \mu\|_{\infty} 
\end{align*}
\\
Using SIP on summation of $k$ terms, we obtain the following upper bound for $|M_1|$
\begin{align*}
|M_1|
    & < 2\|(\bar{\bar{X}} - \mu)\|_{\infty}\sum\limits_{k=-b_n + 1}^{b_n-1}\left[ \dfrac{D \psi(k)}{n} + \dfrac{\|L\| p^{1/2}(1+\epsilon)\sqrt{2k \log\log k}}{n}  \right] \\
    & \quad + 2(2b_n - 1)\|\bar{\bar{X}} - \mu\|_{\infty}\|\bar{X}_1 - \mu\|_{\infty} \\
    &\leq 2(2b_n - 1)\|(\bar{\bar{X}} - \mu)\|_{\infty} \left[ \dfrac{D \psi(n)}{n} + \dfrac{\|L\| p^{1/2}(1+\epsilon)\sqrt{n \log\log n}}{n}  \right]\\
    & \quad  + 2(2b_n - 1)\|\bar{\bar{X}} - \mu\|_{\infty}\|\bar{X}_1 - \mu\|_{\infty}\,.
\end{align*}
%
We use Lemma~\ref{lemma:consis_1} to upper bound the above term as
\begin{equation} \label{eq:strong_consis_term-1}
    4(2b_n - 1)\left[ \dfrac{D \psi(n)}{n} + \dfrac{\|L\| p^{1/2}(1+\epsilon)\sqrt{n \log\log n}}{n}  \right]^2
\end{equation}
\\
Using the bound on $\psi(n)$ given by \cite{stra:1964}, equation (\ref{eq:strong_consis_term-1}) becomes $\mathcal{O}\left(b_n \log \log n/n\right)$ which converges to 0 as $n \to \infty$. Now we will show that $M_2 \xrightarrow{a.s.} 0$ as $n\to \infty$

\begin{align*}
   |M_2| & = \left|\dfrac{1}{m}\sum\limits_{s=1}^{m}\left\{ \left(\mu - \bar{\bar{X}} \right)_i  \left(\mu - \bar{\bar{X}} \right)_j\sum_{k=-b_n+1}^{b_n-1}\left(1-\dfrac{|k|}{n}\right)w\left(\dfrac{k}{b_n}\right)\right\}\right|\\
    & \leq\|\bar{\bar{X}} - \mu\|_{\infty}^2\left[\sum_{k=-b_n+1}^{b_n-1}\left(1-\dfrac{|k|}{n}\right)w\left(\dfrac{k}{b_n}\right)\right]\\
    &< \|\bar{\bar{X}} - \mu\|_{\infty}^2\left[\sum_{k=-b_n+1}^{b_n-1}\left|w\left(\dfrac{k}{b_n}\right)\right|\right]\\
    &= b_n\|\bar{\bar{X}} - \mu\|_{\infty}^2\left[\dfrac{1}{b_n}\sum_{k=-b_n+1}^{b_n-1}\left|w\left(\dfrac{k}{b_n}\right)\right|\right]\\
    & \leq b_n\|\bar{\bar{X}} - \mu\|_{\infty}^2 \int_{-\infty}^{\infty}|w(x)|dx \\
    & \leq Cb_n\|\bar{\bar{X}} - \mu \|^2
\end{align*}

\begin{equation} \label{eq:strong_consis_term-2}
     Cb_n\left[ \dfrac{D \psi(n)}{n} + \dfrac{\|L\| p^{1/2}(1+\epsilon)\sqrt{n \log\log n}}{n}  \right]^2
% Equation~\eqref{eq:strong_consis_term-2} has the same order as equation \eqref{eq:strong_consis_term-1}, i.e. $\mathcal{O}\left(b_n \log \log n/n\right) \to 0 \textrm{ as } n \to \infty$.
